---
title: "`r params$title`"
subtitle: "`r params$subtitle`"
date-format: "YYYY-MM-DD"
date: "`r params$val_date`"
params:
  title: "val.pipeline Qualification Summary"
  subtitle: ""
  qual_metadata_path: NULL
  qual_assessments_path: NULL
  val_date: NULL
  config_path: NULL
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    theme: [cosmo]
    df-print: paged
  typst:
    toc: true
    toc-depth: 3
    section-numbering: 1.1.1
    margin:
      x: 2cm
      y: 2cm
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
options(knitr.kable.NA = "")
library(dplyr)
library(knitr)
```

```{r load-inputs}
qual_metadata <- readRDS(params$qual_metadata_path)
qual_assessments <- if (!is.null(params$qual_assessments_path)) {
  readRDS(params$qual_assessments_path)
} else {
  NULL
}

has_assessments <- !is.null(qual_assessments)

# Backfill columns that older qual_metadata.rds runs may lack (the
# `_note` columns were added in val.pipeline 0.0.1.9003 / issue #37).
for (col in c("decision_reason_note", "final_decision_reason_note")) {
  if (!col %in% names(qual_metadata)) {
    qual_metadata[[col]] <- NA_character_
  }
}
has_notes <- any(!is.na(qual_metadata$final_decision_reason_note))

decision_levels <- c("Low", "Medium", "High")
qual_metadata$final_decision <- factor(qual_metadata$final_decision,
                                       levels = decision_levels)
qual_metadata$decision <- factor(qual_metadata$decision,
                                 levels = decision_levels)
```

# Run Metadata

```{r run-metadata}
r_ver <- unique(qual_metadata$r_ver)
val_dates <- unique(as.character(qual_metadata$val_date))
repos <- unique(unlist(strsplit(as.character(qual_metadata$repos), "\\s*,\\s*")))
metric_pkg <- unique(qual_metadata$metric_pkg)
ref_types <- unique(qual_metadata$ref)

meta_tbl <- tibble::tribble(
  ~Item, ~Value,
  "Report generated",       format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"),
  "Validation date(s)",     paste(val_dates, collapse = ", "),
  "R version",              paste(r_ver, collapse = ", "),
  "Assessment source",      paste(ref_types, collapse = ", "),
  "Metric package",         paste(metric_pkg, collapse = ", "),
  "Repositories",           paste(repos, collapse = "; "),
  "Packages assessed",      format(nrow(qual_metadata), big.mark = ","),
  "qual_metadata path",     params$qual_metadata_path,
  "qual_assessments path",  ifelse(has_assessments, params$qual_assessments_path, "(not provided)")
)
kable(meta_tbl)
```

# Decision Summary

## Final decision counts

The `final_decision` column reflects the risk category after dependency
propagation (a package is downgraded if any dependency was Medium/High).

```{r final-decision-counts}
final_counts <- qual_metadata |>
  dplyr::count(final_decision, .drop = FALSE, name = "packages") |>
  dplyr::mutate(percent = sprintf("%.1f%%", 100 * packages / sum(packages)))
kable(final_counts, col.names = c("Final decision", "Packages", "Percent"),
      align = "lrr")
```

## Initial vs. final decision

Packages whose category changed because of dependency propagation.

```{r decision-crosstab}
xt <- qual_metadata |>
  dplyr::count(decision, final_decision, .drop = FALSE) |>
  tidyr::pivot_wider(names_from = final_decision, values_from = n,
                     values_fill = 0L, names_expand = TRUE)
kable(xt, caption = "Rows = initial decision, Columns = final decision")

downgraded_n <- sum(qual_metadata$decision != qual_metadata$final_decision,
                    na.rm = TRUE)
```

**`r downgraded_n`** package(s) were downgraded from their initial decision
because of a failing (Medium/High) dependency.

## Reasons driving the final decision

```{r decision-reason-counts}
reason_counts <- qual_metadata |>
  dplyr::count(final_decision, final_decision_reason, name = "packages",
               .drop = FALSE) |>
  dplyr::filter(packages > 0) |>
  dplyr::arrange(final_decision, dplyr::desc(packages))
kable(reason_counts,
      col.names = c("Final decision", "Reason", "Packages"),
      align = "llr")
```

# Packages by Risk Category

The tables below list every package in each risk tier, with the reason and
supporting note (driver metrics for `Risk Assessment`, failing dependency
for `Dependency`, auto-accept metric list for `Auto-Accepted`).

```{r pkg-tables-helper}
pkg_table <- function(df, cat) {
  out <- df |>
    dplyr::filter(final_decision == cat) |>
    dplyr::arrange(pkg) |>
    dplyr::transmute(
      Package = pkg,
      Version = ver,
      Reason = final_decision_reason,
      Note = ifelse(is.na(final_decision_reason_note) |
                      final_decision_reason_note == "",
                    "", final_decision_reason_note)
    )
  if (nrow(out) == 0L) {
    return(tibble::tibble(Package = character(),
                          Version = character(),
                          Reason = character(),
                          Note = character()))
  }
  out
}
```

## Low risk (`r sum(qual_metadata$final_decision == "Low", na.rm = TRUE)`)

```{r low-tbl}
kable(pkg_table(qual_metadata, "Low"))
```

## Medium risk (`r sum(qual_metadata$final_decision == "Medium", na.rm = TRUE)`)

```{r medium-tbl}
kable(pkg_table(qual_metadata, "Medium"))
```

## High risk (`r sum(qual_metadata$final_decision == "High", na.rm = TRUE)`)

```{r high-tbl}
kable(pkg_table(qual_metadata, "High"))
```

# Runtime

```{r runtime-summary}
rt <- qual_metadata$assessment_runtime_mins
if (inherits(rt, "difftime")) rt <- as.numeric(rt)
rt <- rt[!is.na(rt)]

if (length(rt) > 0L) {
  rt_tbl <- tibble::tribble(
    ~Metric, ~`Minutes`,
    "Packages with recorded runtime", length(rt),
    "Total runtime (mins)",           round(sum(rt), 1),
    "Total runtime (hours)",          round(sum(rt) / 60, 2),
    "Mean per package (mins)",        round(mean(rt), 3),
    "Median per package (mins)",      round(stats::median(rt), 3),
    "Max per package (mins)",         round(max(rt), 2)
  )
  kable(rt_tbl, align = "lr")
} else {
  cat("No runtime data recorded.\n")
}
```

## Ten slowest packages

```{r slowest-pkgs}
if (length(rt) > 0L) {
  slow_df <- qual_metadata |>
    dplyr::mutate(runtime_mins = as.numeric(assessment_runtime_mins)) |>
    dplyr::filter(!is.na(runtime_mins)) |>
    dplyr::arrange(dplyr::desc(runtime_mins)) |>
    dplyr::slice_head(n = 10) |>
    dplyr::transmute(
      Package = pkg,
      Version = ver,
      `Final decision` = as.character(final_decision),
      `Runtime (mins)` = round(runtime_mins, 2)
    )
  kable(slow_df, align = "llrr")
}
```

```{r assessments-header, results = "asis", eval = has_assessments}
cat("# Metric-Level Summaries\n\n",
    "The following summaries are derived from `qual_assessments.rds` ",
    "and describe the distribution of each `{riskmetric}` metric across ",
    "the assessed packages.\n", sep = "")
```

```{r metric-helpers, eval = has_assessments}
numeric_metric_summary <- function(df, col) {
  x <- suppressWarnings(as.numeric(df[[col]]))
  x <- x[!is.na(x)]
  if (length(x) == 0L) return(NULL)
  tibble::tibble(
    Metric = col,
    N = length(x),
    Min = signif(min(x), 4),
    Median = signif(stats::median(x), 4),
    Mean = signif(mean(x), 4),
    Max = signif(max(x), 4)
  )
}

qa <- qual_assessments
numeric_cols <- vapply(names(qa), function(nm) {
  v <- qa[[nm]]
  if (is.list(v)) return(FALSE)
  suppressWarnings(nv <- as.numeric(v))
  !all(is.na(nv)) && !nm %in% c("val_date", "version")
}, logical(1))
numeric_col_names <- setdiff(
  names(qa)[numeric_cols],
  c("package", "version", "val_date", "riskmetric_version", "ref")
)
```

## Numeric metric distributions

```{r numeric-summaries, eval = has_assessments}
tbl <- purrr::map_dfr(numeric_col_names, ~ numeric_metric_summary(qa, .x))
kable(tbl, align = "lrrrrr")
```

## Package coverage (`covr_coverage`)

```{r coverage, eval = has_assessments && "covr_coverage" %in% names(qa)}
cov <- suppressWarnings(as.numeric(qa$covr_coverage))
cov_tbl <- tibble::tibble(
  Bucket = c("Not measured (NA)", "0%", "1-49%", "50-79%", "80-94%", "95-100%"),
  Packages = c(
    sum(is.na(cov)),
    sum(cov == 0, na.rm = TRUE),
    sum(cov > 0 & cov < 50, na.rm = TRUE),
    sum(cov >= 50 & cov < 80, na.rm = TRUE),
    sum(cov >= 80 & cov < 95, na.rm = TRUE),
    sum(cov >= 95 & cov <= 100, na.rm = TRUE)
  )
) |>
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * Packages / sum(Packages)))
kable(cov_tbl, align = "lrr")
```

## R CMD check outcomes

```{r rcmd-check, eval = has_assessments && all(c("r_cmd_check_errors","r_cmd_check_warnings") %in% names(qa))}
err <- suppressWarnings(as.numeric(qa$r_cmd_check_errors))
wrn <- suppressWarnings(as.numeric(qa$r_cmd_check_warnings))
rcmd_tbl <- tibble::tibble(
  Outcome = c("Clean (0 errors, 0 warnings)",
              "Warnings only",
              "Errors",
              "Not measured (NA)"),
  Packages = c(
    sum(err == 0 & wrn == 0, na.rm = TRUE),
    sum(err == 0 & wrn > 0, na.rm = TRUE),
    sum(err > 0, na.rm = TRUE),
    sum(is.na(err))
  )
) |>
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * Packages / sum(Packages)))
kable(rcmd_tbl, align = "lrr")
```

## License distribution

```{r license, eval = has_assessments && "license" %in% names(qa)}
lic_tbl <- qa |>
  dplyr::count(license, sort = TRUE, name = "packages") |>
  dplyr::mutate(percent = sprintf("%.1f%%", 100 * packages / sum(packages)))
kable(lic_tbl, col.names = c("License", "Packages", "Percent"),
      align = "lrr")
```

# Appendix: Configuration

```{r config, results = "asis"}
if (!is.null(params$config_path) && file.exists(params$config_path)) {
  cat("Configuration snapshot from the run (from `",
      params$config_path, "`):\n\n", sep = "")
  cat("```yaml\n")
  cat(readLines(params$config_path), sep = "\n")
  cat("\n```\n")
} else {
  cat("_(No `config.yml` alongside inputs; configuration snapshot omitted.)_\n")
}
```
