---
title: "`r params$title`"
subtitle: "`r params$subtitle`"
date-format: "YYYY-MM-DD"
date: "`r params$val_date`"
params:
  title: "val.pipeline Qualification Summary"
  subtitle: ""
  qual_metadata_path: NULL
  qual_assessments_path: NULL
  val_date: NULL
  config_path: NULL
  thresholds_path: NULL
  n_candidates: NULL
  pre_filtered_path: NULL
  pipeline_runtime: NULL
  mem_watchdog_path: NULL
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    theme: [cosmo]
    df-print: paged
  typst:
    toc: true
    toc-depth: 3
    section-numbering: 1.1.1
    margin:
      x: 2cm
      y: 2cm
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
options(knitr.kable.NA = "")
library(dplyr)
library(knitr)

is_html <- isTRUE(knitr::is_html_output())

# Interactive filterable/sortable table for HTML output; plain kable for
# typst/PDF. Kept self-contained (i.e. defined in the qmd rather than
# imported from val.pipeline) so `devtools::load_all()` renders work
# without the package being installed. The equivalent internal helper
# `val.pipeline:::render_summary_table()` mirrors this logic for callers
# that want to reuse it programmatically.
itable <- function(df, ..., searchable = TRUE, filterable = TRUE,
                   default_page_size = 25) {
  if (is_html && nrow(df) > 0L) {
    reactable::reactable(
      df,
      searchable = searchable,
      filterable = filterable,
      sortable = TRUE,
      resizable = TRUE,
      defaultPageSize = default_page_size,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(10, 25, 50, 100),
      striped = TRUE,
      highlight = TRUE,
      compact = TRUE,
      wrap = TRUE,
      ...
    )
  } else {
    knitr::kable(df, ...)
  }
}

# Format a duration (minutes) as "Xh Ym" when >= 1h, "Ym" for
# 1 <= mins < 60, "X.XXm" for < 1 minute. Returns "-" for NA / non-
# finite input so table cells never blow up on missing data. See #130.
format_hm <- function(mins) {
  x <- suppressWarnings(as.numeric(mins))
  if (length(x) != 1L || !is.finite(x) || x < 0) return("-")
  if (x < 1) return(sprintf("%.2f m", x))
  rounded_min <- round(x)
  if (rounded_min < 60) return(sprintf("%d m", as.integer(rounded_min)))
  h <- rounded_min %/% 60
  m <- rounded_min - h * 60
  sprintf("%dh %02dm", as.integer(h), as.integer(m))
}
```

```{r load-inputs}
qual_metadata <- readRDS(params$qual_metadata_path)
qual_assessments <- if (!is.null(params$qual_assessments_path)) {
  readRDS(params$qual_assessments_path)
} else {
  NULL
}
pre_filtered <- if (!is.null(params$pre_filtered_path) &&
                      file.exists(params$pre_filtered_path)) {
  readRDS(params$pre_filtered_path)
} else {
  NULL
}

has_assessments <- !is.null(qual_assessments)
has_pre_filter <- !is.null(pre_filtered) &&
                    all(c("package", "final_risk") %in% names(pre_filtered))

# Backfill columns that older qual_metadata.rds runs may lack (the
# `_note` columns were added in val.pipeline 0.0.1.9003 / issue #37).
for (col in c("decision_reason_note", "final_decision_reason_note")) {
  if (!col %in% names(qual_metadata)) {
    qual_metadata[[col]] <- NA_character_
  }
}
has_notes <- any(!is.na(qual_metadata$final_decision_reason_note))

decision_levels <- c("Low", "Medium", "High")
qual_metadata$final_decision <- factor(qual_metadata$final_decision,
                                       levels = decision_levels)
qual_metadata$decision <- factor(qual_metadata$decision,
                                 levels = decision_levels)

# Normalise pre_filtered final_risk to the same factor levels for consistent
# ordering in tables. `val_categorize()` already returns it as a factor, but
# older RDS files may not.
if (has_pre_filter) {
  pre_filtered$final_risk <- factor(as.character(pre_filtered$final_risk),
                                    levels = decision_levels)
  assessed_pkgs <- unique(qual_metadata$pkg)
  pre_filtered$passed_to_build <- pre_filtered$package %in% assessed_pkgs
}
```

# Run Metadata

```{r run-metadata}
r_ver <- unique(qual_metadata$r_ver)
val_dates <- unique(as.character(qual_metadata$val_date))
# Split, unique, and drop "unknown" (unnamed literal returned by
# `get_repo_origin()` in R/utils.R when the pkg's source URL doesn't
# substring-match any entry in `getOption("repos")` at assessment
# time -- typically dep-skipped bundles whose repo was resolved
# before the option was populated, or pkgs pulled from an ad-hoc
# `Additional_repositories:` / GitHub tarball path). Leaving it in
# the "Repositories" row shows a misleading literal in the top meta
# table. Same NA-filter idea already applied to `ref` / `metric_pkg`
# below. See #130.
repos <- unique(unlist(strsplit(as.character(qual_metadata$repos), "\\s*,\\s*")))
repos <- setdiff(repos, c("unknown", NA_character_, ""))
# Drop NAs before flattening: rows whose packages were categorised via
# dependency propagation (never independently assessed by riskmetric)
# leave `ref` / `metric_pkg` unset, and we don't want a literal "NA" to
# leak into the human-readable metadata table.
metric_pkg <- setdiff(unique(qual_metadata$metric_pkg), NA)
ref_types <- setdiff(unique(qual_metadata$ref), NA)

meta_rows <- list(
  c("Report generated",      format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")),
  c("Validation date(s)",    paste(val_dates, collapse = ", ")),
  c("R version",             paste(r_ver, collapse = ", ")),
  c("Assessment reference",  paste(ref_types, collapse = ", ")),
  c("Metric package",        paste(metric_pkg, collapse = ", ")),
  c("Repositories",          paste(repos, collapse = "; "))
)
# val.pipeline version(s) that produced this run. Sourced from the
# per-pkg `val_pipeline_ver` list-col on qual_metadata (populated by
# val_pkg() as of #130). A resumed run can span multiple versions if
# the operator updated the package between sessions -- show the
# distinct set in insertion order so the reader sees the mix.
if ("val_pipeline_ver" %in% names(qual_metadata)) {
  vp_vers <- unique(stats::na.omit(as.character(qual_metadata$val_pipeline_ver)))
  vp_vers <- vp_vers[nzchar(vp_vers)]
  if (length(vp_vers) > 0L) {
    meta_rows <- c(meta_rows, list(
      c("val.pipeline version(s)", paste(vp_vers, collapse = ", "))
    ))
  }
}
# Cumulative pipeline runtime = sum of per-pkg assessment_runtime_mins
# across the cohort. Each pkg only runs once, so summing correctly
# accounts for time spent across resumed sessions -- unlike the prior
# `params$pipeline_runtime` field which was Sys.time() - val_start of
# only the LAST session. See #130.
if ("assessment_runtime_mins" %in% names(qual_metadata)) {
  rt_all <- suppressWarnings(
    as.numeric(qual_metadata$assessment_runtime_mins))
  rt_all <- rt_all[is.finite(rt_all) & rt_all >= 0]
  if (length(rt_all) > 0L) {
    meta_rows <- c(meta_rows, list(
      c("val_pipeline() runtime (cumulative)",
        format_hm(sum(rt_all)))
    ))
  }
}
if (!is.null(params$n_candidates)) {
  meta_rows <- c(meta_rows, list(
    c("Candidate packages evaluated",
      format(as.integer(params$n_candidates), big.mark = ","))
  ))
}
meta_rows <- c(meta_rows, list(
  c("Packages assessed",     format(nrow(qual_metadata), big.mark = ",")),
  c("qual_metadata path",    params$qual_metadata_path),
  c("qual_assessments path", ifelse(has_assessments,
                                    params$qual_assessments_path,
                                    "(not provided)"))
))
meta_tbl <- tibble::tibble(
  Item  = vapply(meta_rows, `[`, character(1), 1),
  Value = vapply(meta_rows, `[`, character(1), 2)
)
kable(meta_tbl)
```

# Pre-Filter Summary

```{r pre-filter-summary, results = "asis", eval = has_pre_filter}
cat("The 'pre-filter step' runs `riskmetric`'s `remote_reduce` rules over ",
    "the entire candidate universe (using recent remote metrics stored in ",
    "`{riskscore}`) to identify ",
    "packages worth running a full assessment on. Only packages ",
    "categorised as **Low** risk here \u2014 plus any explicitly ",
    "allow-listed in the `pass_primary` config \u2014 proceed to ",
    "`val_build()`.\n\n", sep = "")
```

```{r pre-filter-summary-skip, results = "asis", eval = !has_pre_filter}
cat("_No `pre_filtered_pkg_metrics.rds` was supplied or auto-detected ",
    "alongside `qual_metadata.rds`. Historical runs predating this ",
    "artifact will not include a pre-filter summary._\n\n", sep = "")
```

```{r pre-filter-passfail-hdr, results = "asis", eval = has_pre_filter}
cat("## Pass / drop breakdown\n\n")
```

```{r pre-filter-passfail, eval = has_pre_filter}
n_total <- nrow(pre_filtered)
n_passed <- sum(pre_filtered$passed_to_build)
n_dropped <- n_total - n_passed
pf_pass <- data.frame(
  Item = c("Candidate packages evaluated",
           "Passed to val_build()",
           "Dropped by pre-filter"),
  n = c(n_total, n_passed, n_dropped),
  stringsAsFactors = FALSE
) |>
  dplyr::arrange(dplyr::desc(n)) |>
  dplyr::mutate(Count = format(n, big.mark = ",")) |>
  dplyr::select(Item, Count)
kable(pf_pass, align = "lr")
```

```{r pre-filter-counts-hdr, results = "asis", eval = has_pre_filter}
cat("## Pre-filter risk distribution\n\n")
```

```{r pre-filter-counts, eval = has_pre_filter}
pre_counts <- pre_filtered |>
  dplyr::count(final_risk, .drop = FALSE, name = "n") |>
  dplyr::mutate(
    Percent = sprintf("%.1f%%", 100 * n / sum(n)),
    n = format(n, big.mark = ",")
  ) |>
  dplyr::rename(`Pre-filter risk` = final_risk, Count = n)
kable(pre_counts, align = "lrr")
```

```{r pre-filter-per-metric-hdr, results = "asis", eval = has_pre_filter}
cat("## Per-metric risk distribution\n\n")
cat("For each metric evaluated by the `remote_reduce` rules, the ",
    "share of candidate packages that fell into each risk category. ",
    "This mirrors the summary printed to the console during ",
    "`val_pipeline()`.\n\n", sep = "")
```

```{r pre-filter-per-metric, eval = has_pre_filter}
# Pretty labels for metric names shown in the report. Keys are the raw
# column names on the pre_filtered_pkg_metrics data frame (either a
# secondary `_cat` col or the primary_risk_category surrogate for the
# `remote_reduce` primary metric — downloads_1yr).
metric_labels <- c(
  primary_risk_category = "Downloads (1yr)",
  rev_deps_cat          = "Reverse dependencies",
  n_deps_cat            = "Dependencies",
  n_vig_cat             = "Vignettes",
  news_curr_cat         = "NEWS currency",
  src_cntrl_cat         = "Source control",
  n_sites_cat           = "Website",
  bugs_status_cat       = "Bug tracker status"
)

# Fallback: strip the _cat suffix and title-case each word.
pretty_metric <- function(col) {
  if (col %in% names(metric_labels)) return(unname(metric_labels[[col]]))
  base <- sub("_cat$", "", col)
  base <- gsub("_", " ", base)
  substr(base, 1, 1) <- toupper(substr(base, 1, 1))
  base
}

# Assemble the list of columns to summarise, in the order we want them
# displayed: primary first (if present), then the secondary `_cat` cols
# in the order they appear on the data frame.
cat_cols <- grep("_cat$", names(pre_filtered), value = TRUE)
cat_cols <- setdiff(cat_cols, "final_risk_cat")
summary_cols <- c(
  if ("primary_risk_category" %in% names(pre_filtered)) {
    "primary_risk_category"
  },
  cat_cols
)

if (length(summary_cols) > 0L) {
  per_metric <- lapply(summary_cols, function(cc) {
    vals <- factor(as.character(pre_filtered[[cc]]),
                   levels = decision_levels)
    tab <- table(vals, useNA = "no")
    tot <- sum(tab)
    fmt <- function(n) {
      if (tot == 0L) "0 (0.0%)"
      else sprintf("%s (%.1f%%)", format(n, big.mark = ","),
                   100 * n / tot)
    }
    data.frame(
      Metric = pretty_metric(cc),
      Low    = fmt(tab[["Low"]]),
      Medium = fmt(tab[["Medium"]]),
      High   = fmt(tab[["High"]]),
      stringsAsFactors = FALSE
    )
  })
  per_metric_tbl <- do.call(rbind, per_metric)
  itable(per_metric_tbl, default_page_size = 15)
} else {
  cat("_No per-metric risk-category columns available in ",
      "`pre_filtered_pkg_metrics.rds`._\n")
}
```

```{r pre-filter-dropped-hdr, results = "asis", eval = has_pre_filter}
n_dropped_ap <- sum(!pre_filtered$passed_to_build)
cat("## Packages dropped by pre-filter\n\n")
if (is_html) {
  cat("The ", format(n_dropped_ap, big.mark = ","),
      " packages below were evaluated against the `remote_reduce` rules ",
      "but did not qualify for a full assessment run (either their ",
      "pre-filter risk was above `Low`, or they were absent from the ",
      "`pass_primary` allow-list). Their raw metric values are shown for ",
      "context \u2014 these are the reasons `val_build()` never opened ",
      "them.\n\n", sep = "")
} else {
  cat(format(n_dropped_ap, big.mark = ","),
      " packages were evaluated against the `remote_reduce` rules but did ",
      "not qualify for a full assessment run. The per-package listing is ",
      "omitted from the PDF for length; render the HTML version of this ",
      "report to view the filterable table of dropped candidates.\n\n",
      sep = "")
}
```

```{r pre-filter-dropped-table, eval = has_pre_filter && is_html}
metric_cols <- c(
  Package = "package",
  Version = "version",
  `Pre-filter risk` = "final_risk",
  `Auto-pass` = "auto_pass",
  `Downloads (1yr)` = "dwnlds",
  `Reverse deps` = "rev_deps",
  Dependencies = "n_deps",
  Vignettes = "n_vig",
  `NEWS current` = "news_curr",
  `Source control` = "src_cntrl",
  Website = "n_sites",
  Repo = "repo_name"
)
present <- metric_cols[metric_cols %in% names(pre_filtered)]

dropped_tbl <- pre_filtered |>
  dplyr::filter(!passed_to_build) |>
  dplyr::select(dplyr::all_of(unname(present))) |>
  dplyr::arrange(dplyr::desc(final_risk),
                 if ("dwnlds" %in% names(pre_filtered)) dplyr::desc(dwnlds)
                 else dplyr::desc(package))
names(dropped_tbl) <- names(present)

for (col in c("Downloads (1yr)", "Reverse deps", "Dependencies")) {
  if (col %in% names(dropped_tbl)) {
    dropped_tbl[[col]] <- ifelse(
      is.na(dropped_tbl[[col]]),
      NA_character_,
      format(as.numeric(dropped_tbl[[col]]), big.mark = ",",
             scientific = FALSE, trim = TRUE)
    )
  }
}
itable(dropped_tbl, default_page_size = 25)
```

# Decision Summary

## Final decision counts

The `final_decision` column reflects the risk category after dependency
propagation (a package is downgraded if any dependency was Medium/High).

```{r final-decision-counts}
final_counts <- qual_metadata |>
  dplyr::count(final_decision, .drop = FALSE, name = "packages") |>
  dplyr::mutate(percent = sprintf("%.1f%%", 100 * packages / sum(packages)))
kable(final_counts, col.names = c("Final decision", "Packages", "Percent"),
      align = "lrr")
```

## Initial vs. final decision

Packages whose category changed because of dependency propagation.

```{r decision-crosstab-downgrade-callout, results = "asis"}
downgraded_n <- sum(qual_metadata$decision != qual_metadata$final_decision,
                    na.rm = TRUE)
# Large-font callout above the crosstab so the reader's eye lands on
# the headline number first. See #130.
cat("::: {style=\"font-size: 1.8em; font-weight: 600; margin: 0.5em 0 0.75em;\"}\n",
    format(downgraded_n, big.mark = ","),
    " package(s) downgraded from their initial decision\n",
    ":::\n", sep = "")
```

```{r decision-crosstab}
xt <- qual_metadata |>
  dplyr::count(decision, final_decision, .drop = FALSE) |>
  tidyr::pivot_wider(names_from = final_decision, values_from = n,
                     values_fill = 0L, names_expand = TRUE)
kable(xt, caption = "Rows = initial decision, Columns = final decision")
```

Downgrades happen when a runtime dependency failed its assessment
(Medium/High), so PPM cannot serve the parent package with a
complete install closure.

## Ripple effect: failing packages by dep-downgrade count

For every package that got downgraded because of a failing dep
(`final_decision_reason` is `Dependency` or `Pre-Approved (dep
failed)`), the failing direct dep(s) are already named in
`final_decision_reason_note`. This table inverts that mapping to
show which failing packages had the biggest downstream footprint
— i.e. how many other packages each one dragged down.

```{r ripple-effect}
ripple_src <- qual_metadata |>
  dplyr::filter(final_decision_reason %in%
                  c("Dependency", "Pre-Approved (dep failed)")) |>
  dplyr::filter(!is.na(final_decision_reason_note) &
                  nzchar(final_decision_reason_note)) |>
  dplyr::transmute(
    child_pkg = pkg,
    culprits = strsplit(final_decision_reason_note, "\\s*,\\s*")
  ) |>
  tidyr::unnest_longer(culprits) |>
  dplyr::filter(nzchar(culprits))

if (nrow(ripple_src) > 0L) {
  # Look up each culprit's own final decision so operators can see at
  # a glance whether it errored, was dep-driven itself, or landed
  # High on its own risk assessment. Left join tolerates culprits
  # that aren't in qual_metadata (e.g. a system dep not assessed
  # here) by leaving those cells NA.
  culprit_lookup <- qual_metadata |>
    dplyr::select(culprits = pkg,
                  culprit_decision = final_decision,
                  culprit_reason = final_decision_reason)
  ripple_tbl <- ripple_src |>
    dplyr::count(culprits, name = "downgraded_pkgs") |>
    dplyr::left_join(culprit_lookup, by = "culprits") |>
    dplyr::arrange(dplyr::desc(downgraded_pkgs), culprits) |>
    dplyr::slice_head(n = 100) |>
    dplyr::transmute(
      `Failing package`         = culprits,
      `Downgraded pkgs (count)` = downgraded_pkgs,
      `Its own decision`        = as.character(culprit_decision),
      `Its own reason`          = as.character(culprit_reason)
    )
  itable(ripple_tbl, default_page_size = 25)
} else {
  cat("_No dep-driven downgrades recorded in this run._\n")
}
```

## Reasons driving the final decision

```{r decision-reason-counts}
reason_counts <- qual_metadata |>
  dplyr::count(final_decision, final_decision_reason, name = "packages",
               .drop = FALSE) |>
  dplyr::filter(packages > 0) |>
  dplyr::arrange(final_decision, dplyr::desc(packages)) |>
  dplyr::rename(`Final decision` = final_decision,
                Reason = final_decision_reason,
                Packages = packages)
itable(reason_counts, default_page_size = 10)
```

## Pre-Approved packages that failed dependency propagation

Packages on the config `approved_pkgs` list that had to be downgraded
because at least one runtime dependency failed its assessment. PPM can't
serve these packages until either the failing upstream dep is fixed or the
package is removed from `approved_pkgs`. The note column names the direct
DESCRIPTION-level dep(s) that failed. See #110.

```{r preapproved-dep-failed}
preapp_bad <- qual_metadata |>
  dplyr::filter(final_decision_reason == "Pre-Approved (dep failed)") |>
  dplyr::arrange(pkg) |>
  dplyr::transmute(
    Package = pkg,
    Version = ver,
    `Failing dep(s)` = ifelse(is.na(final_decision_reason_note) |
                                final_decision_reason_note == "",
                              "(unrecorded)", final_decision_reason_note)
  )
```

**`r nrow(preapp_bad)`** Pre-Approved package(s) could not be qualified in
this run.

```{r preapproved-dep-failed-tbl}
if (nrow(preapp_bad) > 0L) {
  itable(preapp_bad, default_page_size = 25)
}
```

## Packages that errored during assessment

Packages whose per-package assessment (`val_pkg()`) threw an error mid-run
(e.g. an internal helper collapsing on an unexpected metric shape, a
corrupt tarball, a covr crash). Rather than cancelling the whole build,
`val_build()` catches the error, assigns the highest risk tier, records
the error message as the decision reason note, and continues with the
next package. See #116.

```{r errored-pkgs}
errored <- qual_metadata |>
  dplyr::filter(final_decision_reason == "Error") |>
  dplyr::arrange(pkg) |>
  dplyr::transmute(
    Package = pkg,
    Version = ver,
    Error = ifelse(is.na(final_decision_reason_note) |
                     final_decision_reason_note == "",
                   "(unrecorded)", final_decision_reason_note)
  )
```

**`r nrow(errored)`** package(s) errored during assessment in this run.

```{r errored-pkgs-tbl}
if (nrow(errored) > 0L) {
  itable(errored, default_page_size = 25)
}
```

## Packages with incomplete assessment

Packages whose `val_decision()` rule ladder produced no category —
typically a `remote_only` package with a shrunken viable-metric set,
or a Bioc package where the initial-metrics whitelist was too thin.
`decision` and `final_decision` are left as `NA` for these packages
rather than silently coerced to a tier; per-package detail lives in
the Appendix. See #124.

```{r incomplete-pkgs}
incomplete <- qual_metadata |>
  dplyr::filter(is.na(final_decision) |
                  final_decision_reason == "Incomplete Assessment") |>
  dplyr::arrange(pkg) |>
  dplyr::transmute(
    Package = pkg,
    Version = ver,
    Source  = if ("repo_name" %in% names(qual_metadata))
                as.character(repo_name) else NA_character_,
    Reason  = ifelse(is.na(final_decision_reason) |
                       final_decision_reason == "",
                     "(unrecorded)", final_decision_reason),
    Note    = ifelse(is.na(final_decision_reason_note) |
                       final_decision_reason_note == "",
                     "(no diagnostic)", final_decision_reason_note)
  )
```

**`r nrow(incomplete)`** package(s) could not be categorized in this run.

```{r incomplete-pkgs-tbl}
if (nrow(incomplete) > 0L) {
  itable(incomplete, default_page_size = 25)
}
```

# Packages

The tables below list every package in each risk tier, with the reason and
supporting note (driver metrics for `Risk Assessment`, failing dependency
for `Dependency`, auto-accept metric list for `Auto-Accepted`). In HTML,
use the search box and per-column filters to narrow down.

```{r pkg-tables-helper}
pkg_table <- function(df, cat) {
  out <- df |>
    dplyr::filter(final_decision == cat) |>
    dplyr::arrange(pkg) |>
    dplyr::transmute(
      Package = pkg,
      Version = ver,
      Reason = final_decision_reason,
      Note = ifelse(is.na(final_decision_reason_note) |
                      final_decision_reason_note == "",
                    "", final_decision_reason_note)
    )
  if (nrow(out) == 0L) {
    return(tibble::tibble(Package = character(),
                          Version = character(),
                          Reason = character(),
                          Note = character()))
  }
  out
}
```

## Low risk (`r sum(qual_metadata$final_decision == "Low", na.rm = TRUE)`)

```{r low-tbl}
itable(pkg_table(qual_metadata, "Low"))
```

## Medium risk (`r sum(qual_metadata$final_decision == "Medium", na.rm = TRUE)`)

```{r medium-tbl}
itable(pkg_table(qual_metadata, "Medium"))
```

## High risk (`r sum(qual_metadata$final_decision == "High", na.rm = TRUE)`)

```{r high-tbl}
itable(pkg_table(qual_metadata, "High"))
```

# Runtime

```{r runtime-summary}
has_runtime <- "assessment_runtime_mins" %in% names(qual_metadata)
rt <- if (has_runtime) qual_metadata$assessment_runtime_mins else numeric(0)
if (inherits(rt, "difftime")) rt <- as.numeric(rt)
rt <- rt[!is.na(rt)]

if (length(rt) > 0L) {
  rt_tbl <- tibble::tribble(
    ~Metric, ~Value,
    "Packages with recorded runtime",
      format(length(rt), big.mark = ","),
    "Cumulative runtime",             format_hm(sum(rt)),
    "Mean per package",               format_hm(mean(rt)),
    "Median per package",             format_hm(stats::median(rt)),
    "Max per package",                format_hm(max(rt))
  )
  kable(rt_tbl, align = "lr")
} else {
  cat("_No `assessment_runtime_mins` recorded on this run._\n")
}
```

## Slowest packages

Sourced as the 50 slowest per-package assessments; use the page-size
control to browse fewer rows at a time.

```{r slowest-pkgs}
if (length(rt) > 0L) {
  slow_df <- qual_metadata |>
    dplyr::mutate(runtime_mins = as.numeric(assessment_runtime_mins)) |>
    dplyr::filter(!is.na(runtime_mins)) |>
    dplyr::arrange(dplyr::desc(runtime_mins)) |>
    dplyr::slice_head(n = 50) |>
    dplyr::transmute(
      Package = pkg,
      Version = ver,
      `Final decision` = as.character(final_decision),
      Runtime = vapply(runtime_mins, format_hm, character(1)),
      `Runtime (mins)` = round(runtime_mins, 2)
    )
  itable(slow_df, default_page_size = 10)
} else {
  cat("_(runtime data not available)_\n")
}
```

## Memory offenders

Sourced as the 50 packages with the largest peak RSS during
assessment; use the page-size control to browse fewer rows at a time.

```{r mem-watchdog-load, include = FALSE}
mw_path <- params$mem_watchdog_path
has_mem_watchdog <- !is.null(mw_path) && nzchar(mw_path) &&
  file.exists(mw_path)
mw_df <- if (has_mem_watchdog) {
  tryCatch(
    utils::read.delim(mw_path, sep = "\t", header = TRUE,
                      stringsAsFactors = FALSE,
                      check.names = FALSE, quote = ""),
    error = function(e) NULL
  )
} else {
  NULL
}
if (!is.null(mw_df) && "peak_rss_mb" %in% names(mw_df)) {
  mw_df$peak_rss_mb <- suppressWarnings(as.numeric(mw_df$peak_rss_mb))
  if ("elapsed_sec" %in% names(mw_df)) {
    mw_df$elapsed_sec <- suppressWarnings(as.numeric(mw_df$elapsed_sec))
  }
}
mw_ready <- !is.null(mw_df) && "peak_rss_mb" %in% names(mw_df) &&
  any(is.finite(mw_df$peak_rss_mb))
```

```{r mem-watchdog-summary, results = "asis", eval = mw_ready}
vals <- mw_df$peak_rss_mb[is.finite(mw_df$peak_rss_mb)]
p50 <- round(as.numeric(stats::quantile(vals, 0.50)), 1)
p95 <- round(as.numeric(stats::quantile(vals, 0.95)), 1)
mx  <- round(max(vals), 1)
cat("Distribution of peak RSS per package (sampled from ",
    "`", basename(mw_path), "`): ",
    "p50 = ", p50, " MB, ",
    "p95 = ", p95, " MB, ",
    "max = ", mx, " MB.\n\n", sep = "")
```

```{r mem-watchdog-table, eval = mw_ready}
mw_top <- mw_df |>
  dplyr::filter(is.finite(peak_rss_mb)) |>
  dplyr::arrange(dplyr::desc(peak_rss_mb)) |>
  dplyr::slice_head(n = 50) |>
  dplyr::transmute(
    Package = pkg,
    Version = if ("version" %in% names(mw_df)) version else NA_character_,
    `Peak RSS (MB)` = round(peak_rss_mb, 1),
    `Elapsed (sec)` = if ("elapsed_sec" %in% names(mw_df))
      round(elapsed_sec, 1) else NA_real_,
    Errored = if ("errored" %in% names(mw_df))
      tolower(as.character(errored)) %in% c("true", "1", "t", "yes")
      else FALSE
  )
itable(mw_top, default_page_size = 25)
```

```{r mem-watchdog-empty, results = "asis", eval = !mw_ready}
cat("_No `mem_watchdog.tsv` recorded for this run ",
    "(either `val_build(mem_watchdog = FALSE)`, a fully cached rerun, ",
    "or the peak-RSS sampler was unavailable on the host)._\n", sep = "")
```

```{r assessments-header, results = "asis", eval = has_assessments}
cat("# Metric-Level Summaries\n\n",
    "The following summaries are derived from `qual_assessments.rds` ",
    "and describe the distribution of each `{riskmetric}` metric across ",
    "the assessed packages.\n", sep = "")
```

```{r metric-helpers, eval = has_assessments}
numeric_metric_summary <- function(df, col) {
  x <- suppressWarnings(as.numeric(df[[col]]))
  x <- x[!is.na(x)]
  if (length(x) == 0L) return(NULL)
  tibble::tibble(
    Metric = col,
    N = length(x),
    Min = signif(min(x), 4),
    Median = signif(stats::median(x), 4),
    Mean = signif(mean(x), 4),
    Max = signif(max(x), 4)
  )
}

qa <- qual_assessments
numeric_cols <- vapply(names(qa), function(nm) {
  v <- qa[[nm]]
  if (is.list(v)) return(FALSE)
  suppressWarnings(nv <- as.numeric(v))
  !all(is.na(nv)) && !nm %in% c("val_date", "version")
}, logical(1))
numeric_col_names <- setdiff(
  names(qa)[numeric_cols],
  c("package", "version", "val_date", "riskmetric_version", "ref")
)
```

## Numeric metric distributions

```{r numeric-summaries, eval = has_assessments}
tbl <- purrr::map_dfr(numeric_col_names, ~ numeric_metric_summary(qa, .x))
itable(tbl, default_page_size = 25)
```

## Package coverage (`covr_coverage`)

The 65% boundary mirrors the Low-tier coverage threshold in
`inst/config.yml` (`decide.rules.covr_coverage`) — packages at or
above 65% pass the coverage rule for the Low tier; below it they
land in Medium or higher unless another rule pulls them down.

```{r coverage, eval = has_assessments && "covr_coverage" %in% names(qa)}
cov <- suppressWarnings(as.numeric(qa$covr_coverage))
cov_tbl <- tibble::tibble(
  Bucket = c("Not measured (NA)", "0%", "1-49%", "50-64%",
             "65-79%", "80-94%", "95-100%"),
  Packages = c(
    sum(is.na(cov)),
    sum(cov == 0, na.rm = TRUE),
    sum(cov > 0 & cov < 50, na.rm = TRUE),
    sum(cov >= 50 & cov < 65, na.rm = TRUE),
    sum(cov >= 65 & cov < 80, na.rm = TRUE),
    sum(cov >= 80 & cov < 95, na.rm = TRUE),
    sum(cov >= 95 & cov <= 100, na.rm = TRUE)
  )
) |>
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * Packages / sum(Packages)))
kable(cov_tbl, align = "lrr")
```

## R CMD check outcomes

```{r rcmd-check, eval = has_assessments && all(c("r_cmd_check_errors","r_cmd_check_warnings") %in% names(qa))}
err <- suppressWarnings(as.numeric(qa$r_cmd_check_errors))
wrn <- suppressWarnings(as.numeric(qa$r_cmd_check_warnings))
rcmd_tbl <- tibble::tibble(
  Outcome = c("Clean (0 errors, 0 warnings)",
              "Warnings only",
              "Errors",
              "Not measured (NA)"),
  Packages = c(
    sum(err == 0 & wrn == 0, na.rm = TRUE),
    sum(err == 0 & wrn > 0, na.rm = TRUE),
    sum(err > 0, na.rm = TRUE),
    sum(is.na(err))
  )
) |>
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * Packages / sum(Packages)))
kable(rcmd_tbl, align = "lrr")
```

## License distribution

```{r license, eval = has_assessments && "license" %in% names(qa)}
lic_tbl <- qa |>
  dplyr::count(license, sort = TRUE, name = "packages") |>
  dplyr::mutate(percent = sprintf("%.1f%%", 100 * packages / sum(packages))) |>
  dplyr::rename(License = license, Packages = packages, Percent = percent)
itable(lic_tbl, default_page_size = 15)
```
# Appendix: Incomplete Assessments

For each package in the "Packages with incomplete assessment" table above,
the diagnostic captured by `val_pkg()` at decision time: which metrics
were viable, what per-metric risk category each one landed on, and the
top-level primary/secondary category that the rule ladder emitted.
Populated for runs post-#124; older runs will show "(not captured)".

```{r incomplete-appendix, results = "asis"}
if ("assessment_gaps" %in% names(qual_metadata)) {
  gaps_rows <- which(
    vapply(qual_metadata$assessment_gaps,
           function(x) !is.null(x) && length(x) > 0L,
           logical(1))
  )
} else {
  gaps_rows <- integer(0)
}

if (length(gaps_rows) == 0L) {
  cat("_(No packages with captured assessment gaps in this run.)_\n")
} else {
  qa_lookup <- if (exists("qa", inherits = FALSE) &&
                     is.data.frame(qa) && "pkg" %in% names(qa)) qa else NULL
  for (i in gaps_rows) {
    pkg_i <- qual_metadata$pkg[i]
    ver_i <- qual_metadata$ver[i]
    g     <- qual_metadata$assessment_gaps[[i]]
    cat(sprintf("## %s (v%s)\n\n", pkg_i, ver_i))
    if (!is.null(g$note) && !is.na(g$note) && nzchar(g$note)) {
      cat(g$note, "\n\n")
    }
    prim_cat_i <- g$primary_risk_category
    if (is.null(prim_cat_i)) prim_cat_i <- NA
    sec_cat_i <- g$secondary_risk_category
    if (is.null(sec_cat_i)) sec_cat_i <- NA
    cat(sprintf("- **Primary risk category**: `%s`\n", prim_cat_i))
    cat(sprintf("- **Secondary risk category**: `%s`\n\n", sec_cat_i))

    vm <- g$viable_metrics
    mc <- g$metric_cats
    if (length(mc) > 0L) {
      metric_tbl <- tibble::tibble(
        Metric   = sub("_cat$", "", names(mc)),
        Category = as.character(unname(mc))
      )
      if (!is.null(qa_lookup)) {
        qa_row <- qa_lookup[qa_lookup$pkg == pkg_i, , drop = FALSE]
        if (nrow(qa_row) >= 1L) {
          metric_tbl$Value <- vapply(metric_tbl$Metric, function(m) {
            if (m %in% names(qa_row)) {
              v <- qa_row[[m]][1]
              if (is.null(v) || (length(v) == 1L && is.na(v))) return("NA")
              format(v)
            } else {
              "(not in qual_assessments)"
            }
          }, character(1))
        }
      }
      cat("**Per-metric outcomes:**\n\n")
      print(kable(metric_tbl, align = "llr"))
      cat("\n")
    } else {
      cat("_(No per-metric categories captured.)_\n\n")
    }

    if (length(vm) > 0L) {
      cat("**Viable metrics (", length(vm), "):** `",
          paste(vm, collapse = "`, `"), "`\n\n", sep = "")
    }
  }
}
```

# Appendix: Metric Thresholds

Human-readable summary of the risk-category thresholds pulled from the
`decide` rule block of `config.yml` at report-generation time. Each row is
one metric; each of the `Low`, `Medium`, `High` columns states the
condition on the metric value that lands a package in that category. When
a metric has an `auto_accept` threshold, meeting it short-circuits the
per-metric evaluation. The `promote_min` config knob is intentionally
omitted.

```{r metric-thresholds}
thresholds_df <- if (!is.null(params$thresholds_path) &&
                     file.exists(params$thresholds_path)) {
  readRDS(params$thresholds_path)
} else {
  NULL
}

if (!is.null(thresholds_df) && nrow(thresholds_df) > 0L) {
  itable(thresholds_df, default_page_size = 25)
} else {
  cat("_(Metric thresholds table unavailable.)_\n")
}
```

# Appendix: Configuration

```{r config, results = "asis"}
if (!is.null(params$config_path) && file.exists(params$config_path)) {
  cat("Configuration snapshot from the run (from `",
      params$config_path, "`):\n\n", sep = "")
  cat("```yaml\n")
  cat(readLines(params$config_path), sep = "\n")
  cat("\n```\n")
} else {
  cat("_(No `config.yml` alongside inputs; configuration snapshot omitted.)_\n")
}
```
