---
title: "`r params$title`"
subtitle: "`r params$subtitle`"
date-format: "YYYY-MM-DD"
date: "`r params$val_date`"
params:
  title: "val.pipeline Qualification Summary"
  subtitle: ""
  qual_metadata_path: NULL
  qual_assessments_path: NULL
  val_date: NULL
  config_path: NULL
  thresholds_path: NULL
  n_candidates: NULL
  pre_filtered_path: NULL
  pipeline_runtime: NULL
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    theme: [cosmo]
    df-print: paged
  typst:
    toc: true
    toc-depth: 3
    section-numbering: 1.1.1
    margin:
      x: 2cm
      y: 2cm
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
options(knitr.kable.NA = "")
library(dplyr)
library(knitr)

is_html <- isTRUE(knitr::is_html_output())

# Interactive filterable/sortable table for HTML output; plain kable for
# typst/PDF. Kept self-contained (i.e. defined in the qmd rather than
# imported from val.pipeline) so `devtools::load_all()` renders work
# without the package being installed. The equivalent internal helper
# `val.pipeline:::render_summary_table()` mirrors this logic for callers
# that want to reuse it programmatically.
itable <- function(df, ..., searchable = TRUE, filterable = TRUE,
                   default_page_size = 25) {
  if (is_html && nrow(df) > 0L) {
    reactable::reactable(
      df,
      searchable = searchable,
      filterable = filterable,
      sortable = TRUE,
      resizable = TRUE,
      defaultPageSize = default_page_size,
      showPageSizeOptions = TRUE,
      pageSizeOptions = c(10, 25, 50, 100),
      striped = TRUE,
      highlight = TRUE,
      compact = TRUE,
      wrap = TRUE,
      ...
    )
  } else {
    knitr::kable(df, ...)
  }
}
```

```{r load-inputs}
qual_metadata <- readRDS(params$qual_metadata_path)
qual_assessments <- if (!is.null(params$qual_assessments_path)) {
  readRDS(params$qual_assessments_path)
} else {
  NULL
}
pre_filtered <- if (!is.null(params$pre_filtered_path) &&
                      file.exists(params$pre_filtered_path)) {
  readRDS(params$pre_filtered_path)
} else {
  NULL
}

has_assessments <- !is.null(qual_assessments)
has_pre_filter <- !is.null(pre_filtered) &&
                    all(c("package", "final_risk") %in% names(pre_filtered))

# Backfill columns that older qual_metadata.rds runs may lack (the
# `_note` columns were added in val.pipeline 0.0.1.9003 / issue #37).
for (col in c("decision_reason_note", "final_decision_reason_note")) {
  if (!col %in% names(qual_metadata)) {
    qual_metadata[[col]] <- NA_character_
  }
}
has_notes <- any(!is.na(qual_metadata$final_decision_reason_note))

decision_levels <- c("Low", "Medium", "High")
qual_metadata$final_decision <- factor(qual_metadata$final_decision,
                                       levels = decision_levels)
qual_metadata$decision <- factor(qual_metadata$decision,
                                 levels = decision_levels)

# Normalise pre_filtered final_risk to the same factor levels for consistent
# ordering in tables. `val_categorize()` already returns it as a factor, but
# older RDS files may not.
if (has_pre_filter) {
  pre_filtered$final_risk <- factor(as.character(pre_filtered$final_risk),
                                    levels = decision_levels)
  assessed_pkgs <- unique(qual_metadata$pkg)
  pre_filtered$passed_to_build <- pre_filtered$package %in% assessed_pkgs
}
```

# Run Metadata

```{r run-metadata}
r_ver <- unique(qual_metadata$r_ver)
val_dates <- unique(as.character(qual_metadata$val_date))
repos <- unique(unlist(strsplit(as.character(qual_metadata$repos), "\\s*,\\s*")))
# Drop NAs before flattening: rows whose packages were categorised via
# dependency propagation (never independently assessed by riskmetric)
# leave `ref` / `metric_pkg` unset, and we don't want a literal "NA" to
# leak into the human-readable metadata table.
metric_pkg <- setdiff(unique(qual_metadata$metric_pkg), NA)
ref_types <- setdiff(unique(qual_metadata$ref), NA)

meta_rows <- list(
  c("Report generated",      format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")),
  c("Validation date(s)",    paste(val_dates, collapse = ", ")),
  c("R version",             paste(r_ver, collapse = ", ")),
  c("Assessment reference",  paste(ref_types, collapse = ", ")),
  c("Metric package",        paste(metric_pkg, collapse = ", ")),
  c("Repositories",          paste(repos, collapse = "; "))
)
if (!is.null(params$pipeline_runtime) &&
      nzchar(as.character(params$pipeline_runtime))) {
  meta_rows <- c(meta_rows, list(
    c("val_pipeline() runtime", as.character(params$pipeline_runtime))
  ))
}
if (!is.null(params$n_candidates)) {
  meta_rows <- c(meta_rows, list(
    c("Candidate packages evaluated",
      format(as.integer(params$n_candidates), big.mark = ","))
  ))
}
meta_rows <- c(meta_rows, list(
  c("Packages assessed",     format(nrow(qual_metadata), big.mark = ",")),
  c("qual_metadata path",    params$qual_metadata_path),
  c("qual_assessments path", ifelse(has_assessments,
                                    params$qual_assessments_path,
                                    "(not provided)"))
))
meta_tbl <- tibble::tibble(
  Item  = vapply(meta_rows, `[`, character(1), 1),
  Value = vapply(meta_rows, `[`, character(1), 2)
)
kable(meta_tbl)
```

# Pre-Filter Summary

```{r pre-filter-summary, results = "asis", eval = has_pre_filter}
cat("The pre-filter step runs `riskmetric`'s `remote_reduce` rules over ",
    "the entire candidate universe (from `{riskscore}`) to identify ",
    "packages worth spending a full assessment on. Only packages ",
    "categorised as **Low** risk here \u2014 plus any explicitly ",
    "allow-listed in the `pass_primary` config \u2014 proceed to ",
    "`val_build()`.\n\n", sep = "")
```

```{r pre-filter-summary-skip, results = "asis", eval = !has_pre_filter}
cat("_No `pre_filtered_pkg_metrics.rds` was supplied or auto-detected ",
    "alongside `qual_metadata.rds`. Historical runs predating this ",
    "artifact will not include a pre-filter summary._\n\n", sep = "")
```

```{r pre-filter-passfail-hdr, results = "asis", eval = has_pre_filter}
cat("## Pass / drop breakdown\n\n")
```

```{r pre-filter-passfail, eval = has_pre_filter}
n_total <- nrow(pre_filtered)
n_passed <- sum(pre_filtered$passed_to_build)
n_dropped <- n_total - n_passed
pf_pass <- data.frame(
  Item = c("Candidate packages evaluated",
           "Passed to val_build()",
           "Dropped by pre-filter"),
  n = c(n_total, n_passed, n_dropped),
  stringsAsFactors = FALSE
) |>
  dplyr::arrange(dplyr::desc(n)) |>
  dplyr::mutate(Count = format(n, big.mark = ",")) |>
  dplyr::select(Item, Count)
kable(pf_pass, align = "lr")
```

```{r pre-filter-counts-hdr, results = "asis", eval = has_pre_filter}
cat("## Pre-filter risk distribution\n\n")
```

```{r pre-filter-counts, eval = has_pre_filter}
pre_counts <- pre_filtered |>
  dplyr::count(final_risk, .drop = FALSE, name = "n") |>
  dplyr::mutate(
    Percent = sprintf("%.1f%%", 100 * n / sum(n)),
    n = format(n, big.mark = ",")
  ) |>
  dplyr::rename(`Pre-filter risk` = final_risk, Count = n)
kable(pre_counts, align = "lrr")
```

```{r pre-filter-per-metric-hdr, results = "asis", eval = has_pre_filter}
cat("## Per-metric risk distribution\n\n")
cat("For each metric evaluated by the `remote_reduce` rules, the ",
    "share of candidate packages that fell into each risk category. ",
    "This mirrors the summary printed to the console during ",
    "`val_pipeline()`.\n\n", sep = "")
```

```{r pre-filter-per-metric, eval = has_pre_filter}
# Pretty labels for metric names shown in the report. Keys are the raw
# column names on the pre_filtered_pkg_metrics data frame (either a
# secondary `_cat` col or the primary_risk_category surrogate for the
# `remote_reduce` primary metric — downloads_1yr).
metric_labels <- c(
  primary_risk_category = "Downloads (1yr)",
  rev_deps_cat          = "Reverse dependencies",
  n_deps_cat            = "Dependencies",
  n_vig_cat             = "Vignettes",
  news_curr_cat         = "NEWS currency",
  src_cntrl_cat         = "Source control",
  n_sites_cat           = "Website",
  bugs_status_cat       = "Bug tracker status"
)

# Fallback: strip the _cat suffix and title-case each word.
pretty_metric <- function(col) {
  if (col %in% names(metric_labels)) return(unname(metric_labels[[col]]))
  base <- sub("_cat$", "", col)
  base <- gsub("_", " ", base)
  substr(base, 1, 1) <- toupper(substr(base, 1, 1))
  base
}

# Assemble the list of columns to summarise, in the order we want them
# displayed: primary first (if present), then the secondary `_cat` cols
# in the order they appear on the data frame.
cat_cols <- grep("_cat$", names(pre_filtered), value = TRUE)
cat_cols <- setdiff(cat_cols, "final_risk_cat")
summary_cols <- c(
  if ("primary_risk_category" %in% names(pre_filtered)) {
    "primary_risk_category"
  },
  cat_cols
)

if (length(summary_cols) > 0L) {
  per_metric <- lapply(summary_cols, function(cc) {
    vals <- factor(as.character(pre_filtered[[cc]]),
                   levels = decision_levels)
    tab <- table(vals, useNA = "no")
    tot <- sum(tab)
    fmt <- function(n) {
      if (tot == 0L) "0 (0.0%)"
      else sprintf("%s (%.1f%%)", format(n, big.mark = ","),
                   100 * n / tot)
    }
    data.frame(
      Metric = pretty_metric(cc),
      Low    = fmt(tab[["Low"]]),
      Medium = fmt(tab[["Medium"]]),
      High   = fmt(tab[["High"]]),
      stringsAsFactors = FALSE
    )
  })
  per_metric_tbl <- do.call(rbind, per_metric)
  itable(per_metric_tbl, default_page_size = 15)
} else {
  cat("_No per-metric risk-category columns available in ",
      "`pre_filtered_pkg_metrics.rds`._\n")
}
```

```{r pre-filter-dropped-hdr, results = "asis", eval = has_pre_filter}
n_dropped_ap <- sum(!pre_filtered$passed_to_build)
cat("## Packages dropped by pre-filter\n\n")
if (is_html) {
  cat("The ", format(n_dropped_ap, big.mark = ","),
      " packages below were evaluated against the `remote_reduce` rules ",
      "but did not qualify for a full assessment run (either their ",
      "pre-filter risk was above `Low`, or they were absent from the ",
      "`pass_primary` allow-list). Their raw metric values are shown for ",
      "context \u2014 these are the reasons `val_build()` never opened ",
      "them.\n\n", sep = "")
} else {
  cat(format(n_dropped_ap, big.mark = ","),
      " packages were evaluated against the `remote_reduce` rules but did ",
      "not qualify for a full assessment run. The per-package listing is ",
      "omitted from the PDF for length; render the HTML version of this ",
      "report to view the filterable table of dropped candidates.\n\n",
      sep = "")
}
```

```{r pre-filter-dropped-table, eval = has_pre_filter && is_html}
metric_cols <- c(
  Package = "package",
  Version = "version",
  `Pre-filter risk` = "final_risk",
  `Auto-pass` = "auto_pass",
  `Downloads (1yr)` = "dwnlds",
  `Reverse deps` = "rev_deps",
  Dependencies = "n_deps",
  Vignettes = "n_vig",
  `NEWS current` = "news_curr",
  `Source control` = "src_cntrl",
  Website = "n_sites",
  Repo = "repo_name"
)
present <- metric_cols[metric_cols %in% names(pre_filtered)]

dropped_tbl <- pre_filtered |>
  dplyr::filter(!passed_to_build) |>
  dplyr::select(dplyr::all_of(unname(present))) |>
  dplyr::arrange(dplyr::desc(final_risk),
                 if ("dwnlds" %in% names(pre_filtered)) dplyr::desc(dwnlds)
                 else dplyr::desc(package))
names(dropped_tbl) <- names(present)

for (col in c("Downloads (1yr)", "Reverse deps", "Dependencies")) {
  if (col %in% names(dropped_tbl)) {
    dropped_tbl[[col]] <- ifelse(
      is.na(dropped_tbl[[col]]),
      NA_character_,
      format(as.numeric(dropped_tbl[[col]]), big.mark = ",",
             scientific = FALSE, trim = TRUE)
    )
  }
}
itable(dropped_tbl, default_page_size = 25)
```

# Decision Summary

## Final decision counts

The `final_decision` column reflects the risk category after dependency
propagation (a package is downgraded if any dependency was Medium/High).

```{r final-decision-counts}
final_counts <- qual_metadata |>
  dplyr::count(final_decision, .drop = FALSE, name = "packages") |>
  dplyr::mutate(percent = sprintf("%.1f%%", 100 * packages / sum(packages)))
kable(final_counts, col.names = c("Final decision", "Packages", "Percent"),
      align = "lrr")
```

## Initial vs. final decision

Packages whose category changed because of dependency propagation.

```{r decision-crosstab}
xt <- qual_metadata |>
  dplyr::count(decision, final_decision, .drop = FALSE) |>
  tidyr::pivot_wider(names_from = final_decision, values_from = n,
                     values_fill = 0L, names_expand = TRUE)
kable(xt, caption = "Rows = initial decision, Columns = final decision")

downgraded_n <- sum(qual_metadata$decision != qual_metadata$final_decision,
                    na.rm = TRUE)
```

**`r downgraded_n`** package(s) were downgraded from their initial decision
because of a failing (Medium/High) dependency.

## Reasons driving the final decision

```{r decision-reason-counts}
reason_counts <- qual_metadata |>
  dplyr::count(final_decision, final_decision_reason, name = "packages",
               .drop = FALSE) |>
  dplyr::filter(packages > 0) |>
  dplyr::arrange(final_decision, dplyr::desc(packages)) |>
  dplyr::rename(`Final decision` = final_decision,
                Reason = final_decision_reason,
                Packages = packages)
itable(reason_counts, default_page_size = 10)
```

# Packages by Risk Category

The tables below list every package in each risk tier, with the reason and
supporting note (driver metrics for `Risk Assessment`, failing dependency
for `Dependency`, auto-accept metric list for `Auto-Accepted`). In HTML,
use the search box and per-column filters to narrow down.

```{r pkg-tables-helper}
pkg_table <- function(df, cat) {
  out <- df |>
    dplyr::filter(final_decision == cat) |>
    dplyr::arrange(pkg) |>
    dplyr::transmute(
      Package = pkg,
      Version = ver,
      Reason = final_decision_reason,
      Note = ifelse(is.na(final_decision_reason_note) |
                      final_decision_reason_note == "",
                    "", final_decision_reason_note)
    )
  if (nrow(out) == 0L) {
    return(tibble::tibble(Package = character(),
                          Version = character(),
                          Reason = character(),
                          Note = character()))
  }
  out
}
```

## Low risk (`r sum(qual_metadata$final_decision == "Low", na.rm = TRUE)`)

```{r low-tbl}
itable(pkg_table(qual_metadata, "Low"))
```

## Medium risk (`r sum(qual_metadata$final_decision == "Medium", na.rm = TRUE)`)

```{r medium-tbl}
itable(pkg_table(qual_metadata, "Medium"))
```

## High risk (`r sum(qual_metadata$final_decision == "High", na.rm = TRUE)`)

```{r high-tbl}
itable(pkg_table(qual_metadata, "High"))
```

# Runtime

```{r runtime-summary}
has_runtime <- "assessment_runtime_mins" %in% names(qual_metadata)
rt <- if (has_runtime) qual_metadata$assessment_runtime_mins else numeric(0)
if (inherits(rt, "difftime")) rt <- as.numeric(rt)
rt <- rt[!is.na(rt)]

if (length(rt) > 0L) {
  rt_tbl <- tibble::tribble(
    ~Metric, ~`Minutes`,
    "Packages with recorded runtime", length(rt),
    "Total runtime (mins)",           round(sum(rt), 1),
    "Total runtime (hours)",          round(sum(rt) / 60, 2),
    "Mean per package (mins)",        round(mean(rt), 3),
    "Median per package (mins)",      round(stats::median(rt), 3),
    "Max per package (mins)",         round(max(rt), 2)
  )
  kable(rt_tbl, align = "lr")
} else {
  cat("_No `assessment_runtime_mins` recorded on this run._\n")
}
```

## Ten slowest packages

```{r slowest-pkgs}
if (length(rt) > 0L) {
  slow_df <- qual_metadata |>
    dplyr::mutate(runtime_mins = as.numeric(assessment_runtime_mins)) |>
    dplyr::filter(!is.na(runtime_mins)) |>
    dplyr::arrange(dplyr::desc(runtime_mins)) |>
    dplyr::slice_head(n = 10) |>
    dplyr::transmute(
      Package = pkg,
      Version = ver,
      `Final decision` = as.character(final_decision),
      `Runtime (mins)` = round(runtime_mins, 2)
    )
  kable(slow_df, align = "llrr")
} else {
  cat("_(runtime data not available)_\n")
}
```

```{r assessments-header, results = "asis", eval = has_assessments}
cat("# Metric-Level Summaries\n\n",
    "The following summaries are derived from `qual_assessments.rds` ",
    "and describe the distribution of each `{riskmetric}` metric across ",
    "the assessed packages.\n", sep = "")
```

```{r metric-helpers, eval = has_assessments}
numeric_metric_summary <- function(df, col) {
  x <- suppressWarnings(as.numeric(df[[col]]))
  x <- x[!is.na(x)]
  if (length(x) == 0L) return(NULL)
  tibble::tibble(
    Metric = col,
    N = length(x),
    Min = signif(min(x), 4),
    Median = signif(stats::median(x), 4),
    Mean = signif(mean(x), 4),
    Max = signif(max(x), 4)
  )
}

qa <- qual_assessments
numeric_cols <- vapply(names(qa), function(nm) {
  v <- qa[[nm]]
  if (is.list(v)) return(FALSE)
  suppressWarnings(nv <- as.numeric(v))
  !all(is.na(nv)) && !nm %in% c("val_date", "version")
}, logical(1))
numeric_col_names <- setdiff(
  names(qa)[numeric_cols],
  c("package", "version", "val_date", "riskmetric_version", "ref")
)
```

## Numeric metric distributions

```{r numeric-summaries, eval = has_assessments}
tbl <- purrr::map_dfr(numeric_col_names, ~ numeric_metric_summary(qa, .x))
itable(tbl, default_page_size = 25)
```

## Package coverage (`covr_coverage`)

```{r coverage, eval = has_assessments && "covr_coverage" %in% names(qa)}
cov <- suppressWarnings(as.numeric(qa$covr_coverage))
cov_tbl <- tibble::tibble(
  Bucket = c("Not measured (NA)", "0%", "1-49%", "50-79%", "80-94%", "95-100%"),
  Packages = c(
    sum(is.na(cov)),
    sum(cov == 0, na.rm = TRUE),
    sum(cov > 0 & cov < 50, na.rm = TRUE),
    sum(cov >= 50 & cov < 80, na.rm = TRUE),
    sum(cov >= 80 & cov < 95, na.rm = TRUE),
    sum(cov >= 95 & cov <= 100, na.rm = TRUE)
  )
) |>
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * Packages / sum(Packages)))
kable(cov_tbl, align = "lrr")
```

## R CMD check outcomes

```{r rcmd-check, eval = has_assessments && all(c("r_cmd_check_errors","r_cmd_check_warnings") %in% names(qa))}
err <- suppressWarnings(as.numeric(qa$r_cmd_check_errors))
wrn <- suppressWarnings(as.numeric(qa$r_cmd_check_warnings))
rcmd_tbl <- tibble::tibble(
  Outcome = c("Clean (0 errors, 0 warnings)",
              "Warnings only",
              "Errors",
              "Not measured (NA)"),
  Packages = c(
    sum(err == 0 & wrn == 0, na.rm = TRUE),
    sum(err == 0 & wrn > 0, na.rm = TRUE),
    sum(err > 0, na.rm = TRUE),
    sum(is.na(err))
  )
) |>
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * Packages / sum(Packages)))
kable(rcmd_tbl, align = "lrr")
```

## License distribution

```{r license, eval = has_assessments && "license" %in% names(qa)}
lic_tbl <- qa |>
  dplyr::count(license, sort = TRUE, name = "packages") |>
  dplyr::mutate(percent = sprintf("%.1f%%", 100 * packages / sum(packages))) |>
  dplyr::rename(License = license, Packages = packages, Percent = percent)
itable(lic_tbl, default_page_size = 15)
```

# Appendix: Metric Thresholds

Human-readable summary of the risk-category thresholds pulled from the
`decide` rule block of `config.yml` at report-generation time. Each row is
one metric; each of the `Low`, `Medium`, `High` columns states the
condition on the metric value that lands a package in that category. When
a metric has an `auto_accept` threshold, meeting it short-circuits the
per-metric evaluation. The `promote_min` config knob is intentionally
omitted.

```{r metric-thresholds}
thresholds_df <- if (!is.null(params$thresholds_path) &&
                     file.exists(params$thresholds_path)) {
  readRDS(params$thresholds_path)
} else {
  NULL
}

if (!is.null(thresholds_df) && nrow(thresholds_df) > 0L) {
  itable(thresholds_df, default_page_size = 25)
} else {
  cat("_(Metric thresholds table unavailable.)_\n")
}
```

# Appendix: Configuration

```{r config, results = "asis"}
if (!is.null(params$config_path) && file.exists(params$config_path)) {
  cat("Configuration snapshot from the run (from `",
      params$config_path, "`):\n\n", sep = "")
  cat("```yaml\n")
  cat(readLines(params$config_path), sep = "\n")
  cat("\n```\n")
} else {
  cat("_(No `config.yml` alongside inputs; configuration snapshot omitted.)_\n")
}
```
