---
title: "`r basename(params$package_name)` v`r params$package_version`"
published-title: ""
params:
  package_name: dplyr
  package_version: 1.0.0
  package: NULL
  image: "rhub/ref-image"
  assessment_path: "assessments/dplyr.rds"
  hide_reverse_deps: true
  source: "pkg_install"
  val_date: NULL
  val_dir: NULL
format:
  html: 
    toc: true  
    embed-resources: true
    include-in-header: "top_page.html"
    include-after-body: "transform_pre_class.html"
    theme: 
      dark: [darkly, custom.scss]
      light: [default, custom.scss, custom_light.scss]
  gfm:
    html-math-method: webtex    
  typst:
    toc: true
    section-numbering: 1.1.1
    df-print: default
    margin:
      x: 2cm
      y: 2cm
filters:
  - foldable_code.lua
---

```{r setup, include = FALSE}
options(width = 80L, covr.record_tests = TRUE)

knitr::opts_chunk$set(
    echo = FALSE,
    eval = TRUE,
    error = TRUE,
    cache = FALSE
)
library("tools")
library("knitr")
library("riskreports")
```

# Context


```{r loading}
pkg <- params$package
pkg_name <- params$package_name
pkg_version <- params$package_version
risk_path <- params$assessment_path
image <- params$image
val_date <- params$val_date
val_dir <- params$val_dir

d_riskmetric <- readRDS(risk_path)

# Required fields
required_fields <- c("r_cmd_check", "license", "remote_checks", "covr_coverage")

# Missing fields
missing_fields <- setdiff(required_fields, names(d_riskmetric))

# Assign default error structure for missing fields to avoid errors
for (field in missing_fields) {
  d_riskmetric[[field]] <- structure(NA, class = c("pkg_metric_error", "pkg_metric"))
}
```

```{r context, results='asis'}
cat("This report was generated by the `val.pipeline` package as part of a",
    "package validation run.\n\n", sep = " ")

cat("- **Validation date (`val_date`):** ",
    if (!is.null(val_date) && nzchar(val_date)) val_date else "unknown",
    "\n", sep = "")
cat("- **Validation directory (`val_dir`):** ",
    if (!is.null(val_dir) && nzchar(val_dir)) val_dir else "unknown",
    "\n", sep = "")
```


# Package Assessments

## Metric-based risk assessment

The following metrics are derived from the `riskmetric` R package.

```{r create_r_riskmetric, warning=FALSE}
# Assesment produces a data.frame with one row
r_riskmetric <- riskreports::assessment(d_riskmetric)
is_na <- sapply(r_riskmetric, function(x) {
  is.na(x) || (is.character(x) && startsWith(x, "no applicable"))
})
r_riskmetric$origin <- riskreports:::get_pkg_origin(params$source)
```


::: {.content-visible when-format="html"}

```{r info_cards}
riskreports:::create_metrics_cards(r_riskmetric)
```

:::

::: {.content-hidden when-format="html"}

```{r simple_cards}
safe_scalar <- function(x) {
  if (is.null(x) || length(x) == 0) return("unknown")
  x <- unname(x)
  as.character(x[[1]])
}
rdep_raw <- d_riskmetric$reverse_dependencies
rdep_display <- if (inherits(rdep_raw, "pkg_metric_error") ||
                    is.null(rdep_raw) || length(rdep_raw) == 0) {
  "unknown"
} else if (is.character(rdep_raw)) {
  prettyNum(length(rdep_raw), big.mark = ",")
} else if (is.numeric(rdep_raw)) {
  prettyNum(rdep_raw[[1]], big.mark = ",")
} else {
  "unknown"
}
dl_raw <- r_riskmetric$downloads_1yr
dl_display <- if (is.numeric(dl_raw) && length(dl_raw) >= 1) {
  prettyNum(dl_raw[[1]], big.mark = ",")
} else {
  safe_scalar(dl_raw)
}
simple_cards <- data.frame(
  "Downloads 1 year" = dl_display,
  "Reverse dependencies" = rdep_display,
  "License" = safe_scalar(r_riskmetric$license),
  "Origin" = safe_scalar(r_riskmetric$origin),
  check.names = FALSE,
  row.names = NULL,
  stringsAsFactors = FALSE
)
r_riskmetric$origin <- NULL
knitr::kable(simple_cards)
```

:::

```{r prepare_tables_namespace}
# Some assessments store exported-namespace names with literal surrounding
# quote characters (e.g. cli's `"__cli_update_due"` is 18 bytes long — 16
# for the name plus two `"` bytes). Left as-is, subsetting `export_help`
# by `exported_namespace` returns NA-named entries, which downstream
# triggers `data.frame(... check.names = FALSE): row names contain
# missing values` inside `prepare_namespace_table()`. Strip stray quotes
# from both metrics so the two vectors align.
if (!is.null(d_riskmetric$export_help) &&
    !is.null(names(d_riskmetric$export_help))) {
  names(d_riskmetric$export_help) <- gsub('^"|"$', "",
                                          names(d_riskmetric$export_help))
}
if (!is.null(d_riskmetric$exported_namespace) &&
    is.character(d_riskmetric$exported_namespace)) {
  d_riskmetric$exported_namespace <- gsub('^"|"$', "",
                                          d_riskmetric$exported_namespace)
}
namespace_table <- tryCatch(
  riskreports:::prepare_namespace_table(d_riskmetric),
  error = function(e) {
    htmltools::span(paste0("Exported-namespace table unavailable: ",
                           conditionMessage(e)))
  }
)
```


::: {.content-visible when-format="html"}

```{r prepare_tables_dependencies}
deps <- d_riskmetric$dependencies
deps$package <- gsub("\\n", " ", deps$package)
# row.names(deps) <- NULL
dependencies_table <- htmltools::div(
  htmltools::span(
    "The package has these dependencies:"
  ),
  reactable::reactable(
    deps,
    class = "metrics-table"
  )
)
```

:::

```{r prepare_tables_reverse_dependencies}
rdep_raw <- d_riskmetric$reverse_dependencies
reverse_dependencies <- if (inherits(rdep_raw, "pkg_metric_error") ||
                            is.null(rdep_raw) || length(rdep_raw) == 0) {
  "Reverse dependencies could not be retrieved."
} else {
  paste(as.character(rdep_raw), collapse = ", ")
}
```

```{r prepare_tables_examples}
examples_content <- paste0(
  "There are ",
  if (!is_na["has_examples"]) sum(d_riskmetric$has_examples),
  " examples from ",
  if (!is_na["has_examples"]) length(d_riskmetric$has_examples),
  " help pages (",
  if (!is_na["has_examples"]) {
    sprintf("%.2f%%", sum(d_riskmetric$has_examples) / length(d_riskmetric$has_examples) * 100)
    } else {
      "?"
    },
  ")."
)
```

```{r prepare_tables_news}
news_content <- paste0(
  "The package has ",
  if ("has_news" %in% names(d_riskmetric) && !is.null(d_riskmetric$has_news)) d_riskmetric$has_news,
  " NEWS file and it is ",
  if (is.null(d_riskmetric$news_current) && !d_riskmetric$news_current) "not",
  "current."
)
```

```{r general-riskmetric, warning=FALSE}
#| tab.cap: "**Package general assessment:** Coverage, check results, size, 
#|  download the last year, reverse dependencies and number of dependencies."

# Defensively drop columns whose value is NULL, empty, or entirely NA
# before handing r_riskmetric to summary_table(). This prevents downstream
# data.frame() calls (some of which use check.names = FALSE) from failing
# with "row names contain missing values" when a metric column is an error
# object, a list, or a multi-element vector whose NA-check would recycle.
.keep_cols <- vapply(r_riskmetric, function(x) {
  if (is.null(x) || length(x) == 0) return(FALSE)
  suppressWarnings(!all(is.na(x)))
}, logical(1))
summary_data <- summary_table(r_riskmetric[, .keep_cols, drop = FALSE]) |>
  stats::setNames(c("Metric", "Value"))
row.names(summary_data) <- NULL

# The `summary_table()` helper runs every numeric metric through
# `is_logical()`, which coerces a value of exactly `1` to the string
# `"Yes"`. That reads correctly for boolean flags (has_news, etc.) but
# is misleading for count metrics like `dependencies` and
# `reverse_dependencies` when a package happens to have exactly one.
# Restore the raw counts for those two rows so the reader sees a number.
.restore_count <- function(metric_row_name, count_value) {
  if (is.null(count_value) || length(count_value) == 0) return(invisible())
  n <- suppressWarnings(as.integer(count_value[[1]]))
  if (is.na(n)) return(invisible())
  idx <- which(summary_data$Metric == metric_row_name)
  if (length(idx) == 1L) {
    summary_data$Value[idx] <<- prettyNum(n, big.mark = ",")
  }
}
.restore_count("Dependencies",
               if (inherits(d_riskmetric$dependencies, "data.frame"))
                 nrow(d_riskmetric$dependencies) else
                 length(d_riskmetric$dependencies))
.restore_count("Reverse dependencies",
               if (is.character(d_riskmetric$reverse_dependencies))
                 length(d_riskmetric$reverse_dependencies) else
                 d_riskmetric$reverse_dependencies)
.restore_count("Has vignettes",
               if (is.numeric(d_riskmetric$has_vignettes))
                 d_riskmetric$has_vignettes else NULL)

# --- Extra polish on individual summary rows -----------------------------
.set_value <- function(metric_row_name, val) {
  if (is.null(val) || !length(val) || !nzchar(as.character(val))) {
    return(invisible())
  }
  idx <- which(summary_data$Metric == metric_row_name)
  if (length(idx) == 1L) summary_data$Value[idx] <<- as.character(val)
}

# Export help: show "N / M (P%)" derived from the raw per-export logical.
.eh <- d_riskmetric$export_help
if (is.logical(.eh) && length(.eh) > 0) {
  .n <- sum(.eh, na.rm = TRUE)
  .m <- length(.eh)
  .set_value("Export help",
             sprintf("%s / %s (%.1f%%)",
                     prettyNum(.n, big.mark = ","),
                     prettyNum(.m, big.mark = ","),
                     100 * .n / .m))
}

# Bugs status: round the percent-closed figure to one decimal place.
.bs <- d_riskmetric$bugs_status
if (is.logical(.bs) && length(.bs) > 0) {
  .set_value("Bugs status",
             sprintf("%.1f%% closed", 100 * sum(.bs, na.rm = TRUE) / length(.bs)))
}

# Size codebase: pretty-print with a thousands separator.
.sc <- d_riskmetric$size_codebase
if (is.numeric(.sc) && length(.sc) >= 1) {
  .set_value("Size codebase",
             prettyNum(as.integer(.sc[[1]]), big.mark = ","))
}

# Websites: show the associated URL(s) rather than a boolean "Yes".
.ws <- d_riskmetric$has_website
if (is.character(.ws) && length(.ws) > 0) {
  .ws <- .ws[nzchar(.ws) & !is.na(.ws)]
  if (length(.ws) > 0) {
    .set_value("Has website", paste(.ws, collapse = ", "))
  }
}

# Bug reports URL: the `has_bug_reports_url` metric only records presence
# (0/1), not the URL itself. Try to extract the URL from the package's
# sourced DESCRIPTION file when the run's val_dir is available.
if (!is.null(val_dir) && nzchar(as.character(val_dir))) {
  .desc_path <- file.path(as.character(val_dir), "sourced",
                          as.character(pkg_name), "DESCRIPTION")
  if (file.exists(.desc_path)) {
    .desc <- tryCatch(read.dcf(.desc_path), error = function(e) NULL)
    if (!is.null(.desc)) {
      .br <- if ("BugReports" %in% colnames(.desc)) .desc[1, "BugReports"] else NA
      if (!is.na(.br) && nzchar(.br)) {
        .set_value("Has bug reports url", as.character(.br))
      }
    }
  }
}

# Reorder rows into a friendlier grouping:
#   1. Dependencies (top)
#   2. Has news + News current (paired)
#   3. Exported namespace + Export help (paired)
#   4. Has examples + Has vignettes (paired)
#   5. URL-returning fields: Has website, Has source control, Has bug
#      reports url, followed immediately by Bugs status.
#   6. Everything else (e.g. Size codebase) keeps its original relative
#      order after the groups above.
#   7. Has maintainer (bottom)
.preferred_order <- c(
  "Dependencies",
  "Has news", "News current",
  "Exported namespace", "Export help",
  "Has examples", "Has vignettes",
  "Has website", "Has source control", "Has bug reports url", "Bugs status"
)
.bottom_pinned <- c("Has maintainer")
.metric_order  <- summary_data$Metric
.head_rows     <- intersect(.preferred_order, .metric_order)
.tail_rows     <- intersect(.bottom_pinned, .metric_order)
.middle_rows   <- setdiff(.metric_order, c(.head_rows, .tail_rows))
.new_order     <- c(.head_rows, .middle_rows, .tail_rows)
summary_data   <- summary_data[match(.new_order, summary_data$Metric), , drop = FALSE]
row.names(summary_data) <- NULL

values_collapsed_rows <- c(
  "Has news",
  "Exported namespace",
  "Dependencies",
  "Reverse dependencies",
  "Has examples"
)

indexes_collapsed_rows <- which(summary_data$Section %in% values_collapsed_rows)

is_html <- knitr::is_html_output(excludes = "markdown")

# Certain rows of this table are collapsible, containing detailed information.
if (isTRUE(is_html)) {
  reactable::reactable(
  summary_data,
  rownames = FALSE,
  pagination = FALSE,
  details = function(index) {
    if(index %in% indexes_collapsed_rows) {
      collapsed_content <- switch(summary_data$Section[index],
        "Exported namespace" = namespace_table,
        "Dependencies" =  dependencies_table,
        "Reverse dependencies" = reverse_dependencies,
        "Has examples" = examples_content,
        "Has news" = news_content
      )
      htmltools::div(
        style = "padding: 1rem;",
        collapsed_content
      )
    }
  },
  columns = list(
    Section = reactable::colDef(
      name = ""
    ),
    Values = reactable::colDef(
      name = "",
      cell = function(value) {
        if (value ==  "Yes") {
          # Green 'Yes'
          htmltools::span(value, style = "color:var(--bs-success); font-weight: bold;")
        } else if (value == "No") {
          # Red 'No'
          htmltools::span(value, style = "color: var(--bs-danger); font-weight: bold;")
        } else {
          htmltools::span(value, style = "color: var(--bs-body-color); font-weight: bold;")
        }
      },
      align = "center"
    )
  ),
  class = "metrics-table"
)
} else {
  knitr::kable(summary_data)
}
```

```{r license, eval=FALSE, results='asis'}
# License information already appears in the summary card table above;
# rendering it again as a standalone section is redundant, so this chunk
# is intentionally disabled.
```


```{r code_checks_prep, include=FALSE}
# Extract error / warning message text from d_riskmetric$r_cmd_check for
# display in the Code checks section. Notes are intentionally omitted.
# r_cmd_check can arrive in a few shapes:
#   - a named integer vector of counts (successful run:
#     c(notes = 0, errors = 1, warnings = 0))
#   - a `pkg_metric_error` + `rcmdcheck_error` with $result$errors / $result$warnings
#   - a `pkg_metric_error` + `system_command_status_error` with $stderr
#   - a `pkg_metric_error` + simpleError with just $message
.rcmd <- d_riskmetric[["r_cmd_check"]]
.rcmd_errors   <- character(0)
.rcmd_warnings <- character(0)
.rcmd_err_count  <- 0L
.rcmd_warn_count <- 0L

# Best-effort extraction of the raw counts, whether the metric is a bare
# named vector or nested inside a list-shaped condition.
.rcmd_count_of <- function(x, nm) {
  if (is.null(x)) return(0L)
  cand <- x[[nm]] %||% NULL
  if (is.null(cand) && is.list(x) && !is.null(x[["result"]])) {
    cand <- x[["result"]][[nm]]
  }
  if (is.null(cand)) return(0L)
  # a character vector of messages -> count = length; a numeric -> that count
  if (is.character(cand)) return(length(cand))
  n <- suppressWarnings(as.integer(cand[[1]]))
  if (is.na(n)) 0L else n
}
`%||%` <- function(a, b) if (is.null(a)) b else a

.rcmd_err_count  <- .rcmd_count_of(.rcmd, "errors")
.rcmd_warn_count <- .rcmd_count_of(.rcmd, "warnings")

if (!is.null(.rcmd)) {
  # rcmdcheck-style result
  res <- tryCatch(.rcmd$result, error = function(e) NULL)
  if (!is.null(res)) {
    if (!is.null(res$errors))   .rcmd_errors   <- as.character(res$errors)
    if (!is.null(res$warnings)) .rcmd_warnings <- as.character(res$warnings)
  }
  # Named-list count with a companion text field
  if (is.list(.rcmd) && length(.rcmd_errors) == 0 && !is.null(.rcmd$errors)) {
    e <- .rcmd$errors
    if (is.character(e)) .rcmd_errors <- e
  }
  if (is.list(.rcmd) && length(.rcmd_warnings) == 0 && !is.null(.rcmd$warnings)) {
    w <- .rcmd$warnings
    if (is.character(w)) .rcmd_warnings <- w
  }
  # Bare condition object — surface its message as an error
  if (length(.rcmd_errors) == 0 && length(.rcmd_warnings) == 0 &&
      inherits(.rcmd, "pkg_metric_error")) {
    msg <- tryCatch(conditionMessage(.rcmd), error = function(e) NULL)
    if (is.null(msg) && is.list(.rcmd) && !is.null(.rcmd$message)) msg <- .rcmd$message
    if (is.null(msg) && is.list(.rcmd) && !is.null(.rcmd$stderr))  msg <- .rcmd$stderr
    if (!is.null(msg) && nzchar(msg)) .rcmd_errors <- as.character(msg)
  }
}

.rcmd_errors   <- .rcmd_errors[nzchar(.rcmd_errors)]
.rcmd_warnings <- .rcmd_warnings[nzchar(.rcmd_warnings)]

# When R CMD check reported counts > 0 but the underlying assessment did
# not capture the full text (a common pattern for `pkg_metric_r_cmd_check`
# named-integer results), fall back to a placeholder so the reader isn't
# left wondering why the section is empty.
.rcmd_errors_missing_text <- length(.rcmd_errors) == 0 && .rcmd_err_count > 0
.rcmd_warnings_missing_text <- length(.rcmd_warnings) == 0 && .rcmd_warn_count > 0

.has_code_checks <- !riskreports::is_risk_error(d_riskmetric[["r_cmd_check"]]) ||
  !riskreports::is_risk_error(d_riskmetric[["remote_checks"]]) ||
  length(.rcmd_errors) > 0 || length(.rcmd_warnings) > 0 ||
  .rcmd_err_count > 0 || .rcmd_warn_count > 0
```

```{r code_checks_header, eval=.has_code_checks, results='asis'}
cat("## Code checks", "\n")

cat("Code checks for this package:", "\n")
```

```{r r_cmd_check, eval=!riskreports::is_risk_error(d_riskmetric[["r_cmd_check"]]), results='asis'}
r_cmd <- d_riskmetric[["r_cmd_check"]]
cat("> **R CMD check**: ", paste(stringr::str_to_title(names(r_cmd)), r_cmd, sep = " - "), "\n")
```

```{r r_cmd_check_errors, eval=length(.rcmd_errors) > 0, results='asis'}
cat("\n\n> **R CMD check errors:**\n>\n")
for (msg in .rcmd_errors) {
  cat("> ```\n", paste0("> ", strsplit(msg, "\n", fixed = TRUE)[[1]], collapse = "\n"), "\n> ```\n>\n", sep = "")
}
cat("\n")
```

```{r r_cmd_check_errors_no_text, eval=.rcmd_errors_missing_text, results='asis'}
cat(sprintf(
  "\n\n> * %d error%s reported, but the error text was not captured in the assessment output.\n\n",
  .rcmd_err_count, if (.rcmd_err_count == 1L) "" else "s"
))
```

```{r r_cmd_check_warnings, eval=length(.rcmd_warnings) > 0, results='asis'}
cat("\n\n> **R CMD check warnings:**\n>\n")
for (msg in .rcmd_warnings) {
  cat("> ```\n", paste0("> ", strsplit(msg, "\n", fixed = TRUE)[[1]], collapse = "\n"), "\n> ```\n>\n", sep = "")
}
cat("\n")
```

```{r r_cmd_check_warnings_no_text, eval=.rcmd_warnings_missing_text, results='asis'}
cat(sprintf(
  "\n\n> * %d warning%s reported, but the warning text was not captured in the assessment output.\n\n",
  .rcmd_warn_count, if (.rcmd_warn_count == 1L) "" else "s"
))
```



```{r remote_checks, eval=!riskreports::is_risk_error(d_riskmetric[["remote_checks"]]), results='asis'}
cat("> **Remote checks**: ", d_riskmetric[["remote_checks"]], "\n")
```


```{r dependencies_section_prep, include=FALSE}
.deps_raw <- d_riskmetric$dependencies
.deps_names <- character(0)
if (inherits(.deps_raw, "data.frame") && "package" %in% names(.deps_raw)) {
  .deps_names <- as.character(.deps_raw$package)
} else if (is.character(.deps_raw)) {
  .deps_names <- as.character(.deps_raw)
}
.deps_names <- .deps_names[nzchar(.deps_names) & !is.na(.deps_names)]
.deps_names <- sort(unique(.deps_names))
.has_deps <- length(.deps_names) > 0

.rdep_raw <- d_riskmetric$reverse_dependencies
.rdep_names <- character(0)
if (!inherits(.rdep_raw, "pkg_metric_error") && is.character(.rdep_raw)) {
  .rdep_names <- as.character(.rdep_raw)
}
.rdep_names <- .rdep_names[nzchar(.rdep_names) & !is.na(.rdep_names)]
.rdep_names <- sort(unique(.rdep_names))
.has_rdeps <- length(.rdep_names) > 0
```

```{r dependencies_section, eval=.has_deps, results='asis'}
cat("## Dependencies", "\n\n")
cat(sprintf("This package has %s first-order dependenc%s:\n\n",
            prettyNum(length(.deps_names), big.mark = ","),
            if (length(.deps_names) == 1L) "y" else "ies"))
cat(paste(.deps_names, collapse = ", "), ".\n\n", sep = "")
```

```{r reverse_dependencies_section, eval=.has_rdeps, results='asis'}
cat("## Reverse dependencies", "\n\n")
cat(sprintf("This package has %s reverse dependenc%s:\n\n",
            prettyNum(length(.rdep_names), big.mark = ","),
            if (length(.rdep_names) == 1L) "y" else "ies"))
cat(paste(.rdep_names, collapse = ", "), ".\n\n", sep = "")
```


```{r covr_coverage_eval_flag, include=FALSE}
.covr_val <- d_riskmetric[["covr_coverage"]]
.covr_scalar <- if (is.list(.covr_val)) {
  if (!is.null(.covr_val$totalcoverage)) .covr_val$totalcoverage else .covr_val[[1]]
} else {
  .covr_val
}
.covr_ok <- !riskreports::is_risk_error(.covr_val) &&
  !is.null(.covr_scalar) && length(.covr_scalar) == 1 &&
  !is.na(.covr_scalar) && (is.numeric(.covr_scalar) || is.character(.covr_scalar))
.covr_display <- if (.covr_ok && is.numeric(.covr_scalar)) {
  sprintf("%.2f%%", .covr_scalar)
} else if (.covr_ok) {
  as.character(.covr_scalar)
} else {
  ""
}
```

```{r covr_coverage, eval=.covr_ok, results='asis'}
cat("## Code coverage", "\n")

cat("Code coverage for this package: ", .covr_display)
```


# Installation environment


## System Info

```{r execution_info}
#| tab.cap: "**System information**. Table about the system used to check the package."
tt_sys_info_df <- data.frame(
  Field = c("OS", "Platform", "System", "Execution Time"),
  Value = c(
    sessionInfo()$running,
    R.version$platform,
    R.version$system,
    format(Sys.time(), tz = "UTC", usetz = TRUE)
  ))

knitr::kable(tt_sys_info_df)
```

<!--
From here we use the lua filter based on this issue from Quarto itself
https://github.com/quarto-dev/quarto-cli/issues/341
Please take into account that not all content in the chunks will be correctly folded
- character vectors are
- reactable tables are not
An alternative is to use callout blocks with collapse = TRUE
The advantage of the foldable_code lua filter is that the code is printed exactly as in the R console
-->

## R Session Info

```{r session_info, attr.output='.details summary="Session info"'}
sessionInfo()
```

```{r Platform, attr.output='.details summary="Platform"'}
unlist(.Platform)
```

```{r capabilities, attr.output='.details summary="R\'s capabilities"'}
capabilities()
```

```{r external-software, attr.output='.details summary="External software"'}
extSoftVersion()
```

```{r graphic-software, attr.output='.details summary="Graphics external software"'}
grSoftVersion()
```

```{r machine, attr.output='.details summary="Numerical characteristics of the machine"'}
unlist(.Machine)
```

```{r RNGKind, attr.output='.details summary="Random number generation process"'}
RNGkind()
```

## Information about the environment
Environmental and options variables affect how package checks and software in it might behave.

```{r computing, attr.output='.details summary="Environmental variables when running this report"'}
riskreports::environ_report(exclude = c("GITHUB_PAT", "APPDATA", 
                                        "GITHUB", "GITLAB_PAT", "GITLAB",
                                        "ProgramFiles", "ProgramFiles(x86)",
                                        "R_USER", "USERNAME", "USERPROFILE", 
                                        "USERDNSDOMAIN", "USERDOMAIN", 
                                        "USERDOMAIN_ROAMINGPROFILE", "R_DptShare"              
                                        ))
```

```{r options, attr.output='.details summary="These are the options set to generate the report"'}
riskreports::options_report(exclude = c("usethis.full_name", "usethis.description", "usethis.destdir"))
```
